---
title: "Los_Pedros_Spectus"
author: "Bijan GURUNG"
date: "2023-03-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Extracting xy-coordinates of a polygon file


```{r}
library(sf)
```

## Polygon file or a shapefile

Los Padres National Forest shapefile. 

```{r}
boundary <- st_read("D:/CA_Forest_Project/Los_Padres/Los_Padres.shp")

```

Details about the shapefile using sf package. 

```{r, echo = FALSE}
st_geometry_type(boundary)
st_crs(boundary)
st_bbox(boundary)
boundary
```
Convert the polygon to vertices or points and then save the xy-coordinates as a .csv file. The points can be used in geospatial analysis with other layers and platform. However, the format of xy-coordinates can be a bit tricky and confusing. It could take time to convert the format. So, I recommend to use the ArcGIS Pro geoprocessing tool, which is quite straight forward and saves a lot of your time.  

```{r}
st_coordinates(boundary)

st_cast(boundary, "POINT")

pts <- st_cast(boundary,"POINT")
pts

st_write(pts,"xy_pts.csv",layer_options="GEOMETRY=AS_XY")
```

